---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mlr3fda

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![RCMD Check](https://github.com/mlr-org/mlr3fda/actions/workflows/rcmdcheck.yaml/badge.svg)](https://github.com/mlr-org/mlr3fda/actions/workflows/rcmdcheck.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/mlr3fda)](https://CRAN.R-project.org/package=mlr3fda)
<!-- badges: end -->

The goal of mlr3fda is to extend `mlr3` to
[functional data](https://en.wikipedia.org/wiki/Functional_data_analysis).
We use the `tfd_irreg` datatype that is defined in the [tf](https://github.com/fabian-s/tf) R package.

## Installation

You can install the development version of `mlr3fda` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("mlr-org/mlr3fda")
```

## Example

After loading `mlr3fda` a new feature type is available.

```{r task_feature_types}
library(mlr3)
mlr_reflections$task_feature_types

library(mlr3fda)
mlr_reflections$task_feature_types
```

For tutorials on how to create functional data, see the [documentation](https://github.com/fabian-s/tf)
of the `tf` package.

```{r data, dpi = 300}
data("fuel", package = "mlr3fda")

plot(fuel$UVVIS)
```

Data with `tf_irreg` columns can be converted to a `Task` just like any other
dataset.

```{r task}
task = as_task_regr(fuel, target = "heatan")
task
```

Currently there are no `Learner`s that directly operate on functional data.
However one can extract simple functional features using the `PipeOpFFS`.


```{r ffs, fig.width = 5, fig.height = 3}
library("mlr3fda")
library("mlr3verse")

po_fmean = po("ffs", feature = "mean", window = 5)

po_fmean$train(list(task))[[1L]]
```

This can be combined with a `Learner` into a complete `Graph`.

```{r graph}
library(mlr3pipelines)

ids = partition(task)

# drop = TRUE means we remove the functional columns as features
graph = po("ffs", feature = "mean", window = 5, drop = TRUE) %>>%
  po("learner", learner = lrn("regr.rpart"))

glrn = as_learner(graph)

glrn$train(task, row_ids = ids$train)

glrn$predict(task, row_ids = ids$test)

```


