---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mlr3fda

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![RCMD Check](https://github.com/mlr-org/mlr3fda/actions/workflows/rcmdcheck.yaml/badge.svg)](https://github.com/mlr-org/mlr3fda/actions/workflows/rcmdcheck.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/mlr3fda)](https://CRAN.R-project.org/package=mlr3fda)
<!-- badges: end -->

`mlr3` extension for functional data analysis.


## What is `mlr3fda`

`mlr3fda` extends `mlr3` to [functional data](https://en.wikipedia.org/wiki/Functional_data_analysis).
It introduces introduces the feature type `"functional"`, as well as basic
infrastructure for common operations, such as flattening of functional data
using e.g. `PipeOpFlatFunct`, or feature extraction from functional data via
`PipeOpFFE` (Functional Feature Extraction).

## Installation

```r
remotes::install_github("mlr-org/mlr3fda")
```

## Example Usage

Build a `GraphLearner` that first extracts features from the functional 
data and then fits a standard random forest.

```{r}
library(mlr3fda)
library(mlr3verse)

task = tsk("fuel")
print(task)

ids = partition(task)

# define the features we want to extract
extractors = list(
  mean = extractor_mean(na.rm = TRUE),
  max = extractor_max(na.rm = TRUE),
  slope = extractor_slope()
)
graph = po("ffe", extractors = extractors) %>>%
  po("learner", learner = lrn("regr.ranger"))

glrn = as_learner(graph)

glrn$train(task, row_ids = ids$train)

glrn$predict(task, row_ids = ids$test)
```
